### 2022-02-01 writing/NB-IoT_scoping.tex
Starting the project

\documentclass[aspectratio=43,t]{beamer}
\usetheme{KTH}

% remove this if using XeLaTeX or LuaLaTeX
\usepackage[utf8]{inputenc}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{ragged2e}
\usepackage{outlines}
\usepackage{lipsum}
\usepackage{minted}
\usepackage{tikz}
\usepackage{array}
\usepackage{algorithm,algorithmicx}
\usepackage{algpseudocode}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage[export]{adjustbox}

%\setbeamersize{text margin left=0pttext margin right=0pt}
%\addtobeamertemplate{frametitle}{\vspace*{2cm}}{\vspace*{-2cm}}
\addtobeamertemplate{frametitle}{\vspace*{0pt}}{\vspace*{15pt}}
\setbeamersize{text margin left=45pt, text margin right=15pt}

\begin{document}

%------------------------------------------------
\begin{frame}[noframenumbering,plain]
%  \vspace{0.02\textheight}
  \vspace{0.30\textheight}

\begin{columns}[]
\column{30em}
\Large{\centerline{\usebeamercolor[fg]{title}Spectral efficiency in NB-IoT}}

\vspace{0.1\textheight}

\small{\centerline{Sigurgeir Gunnarsson}}
\scriptsize{\centerline{\tt sgun@kth.se}}
\scriptsize{\centerline{}}
\end{columns}
\end{frame}

%------------------------------------------------
\usebackgroundtemplate{\vbox{\null\vspace{3mm}
  \hspace{3mm}\pgfuseimage{kthlogosmall}\par
  \vspace{72mm}\hbox{\hspace{-75mm}\pgfuseimage{kthplatta}}}}

%------------------------------------------------
\begin{frame}
\frametitle{Assumptions}

The type of NB-IoT is the one where it is coexisting with LTE i.e. the NB-IoT uses the LTE control resource blocks (RB) to transmit [Schlienz, 2016].

~\\

There are number of coverage classes, each with its own transmission time interval (TTI) repetition pattern i.e. how many TTI the UE gets assigned in a given cycle.

~\\

The repetition pattern is fixed for each class and the challenge is to assign a class to the UE's to maximize efficiency.

%The challenge is to optimize transmission time interval (TTI) scheduling of different coverage classes. That means different classes are scheduled to transmit with dynamic pattern like every other TTI for one class or two out of three consecutive TTI for another.

\end{frame}

%------------------------------------------------
%\begin{frame}
%\frametitle{Questions}
%
%\textbf{Questions}
%
%\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Problem statement}

%The coexistence with LTE and the fact that NB-IoT uses the same signalling resources as LTE, limits the capacity of the NB-IoT platform.

NB-IoT uses the same signalling structure as LTE, which is a finite resource in terms of capacity. When the number of UE's increases in the cell, efficiency of the scheduling becomes more important.

~\\

The UE's are located at different distance from the base station (BS) with different radio conditions which they compensate for with having different coverage classes. Each coverage class is assigned a different number of TTIs in each cycle.

~\\

The challenge is to assign the correct class to the UE's to maximize data capacity and number of served UE's.

\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Problem statement \small{cont.}}

The goal of the project is:

\begin{outline}
\1 Establish an understanding of the capacity of NB-IoT
  \2 Show with simulations how close to the theoretical capacity is possible to reach
\1 Apply reinforcement learning algorithm on the scheduler
  \2 The goal is trying to increase the capacity or get closer to the theoretical limit
\end{outline}

\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Topics to be covered}

\begin{outline}
\1 NB-IoT
  \2 Structure and usage
  \2 Coexistence with LTE
\1 Capacity and congestion
  \2 Establish levels where service begins to degrade
\1 Machine learning (ML)
  \2 Reinforcement learning
\end{outline}

\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Work to be done}

\begin{outline}
\1 Simulate usage
  \2 Create simulation with load around the maximum capacity
  \2 Monte carlo ?
\1 Apply ML techniques on the simulation
  \2 Verify if it increases capacity
\end{outline}

\end{frame}

%------------------------------------------------
\begin{frame}

\vspace{3.5cm}
\Huge{\hspace{3.4cm}\usebeamercolor[fg]{title}Notes}

\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Action points regarding scoping of the project}

How to adapt to the traffic for scheduling allocation of shared channel resources

\begin{itemize}\itemsep1em
 \justifying
 \item learning is from the view of the BTS
 \item different classes should have different TTI i.e. how many TTI per class
 \item Use traffic pattern from human2human (H2H) traffic
 \item create simulation data from that
\end{itemize}

\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Thoughts and points}

The NB-IoT radio interface has to be described and put into context of existing systems [use Schlienz, 2016].

~\\

Identify when the load is such that efficiency starts to become an issue [Harwahyu, 2017].

~\\

Work from that result and start making adjustements on the access of different coverage classes.

~\\

Dig further into the problem of having multiple coverage classes. Identify if it is a near far problem or something else [use Azari. 2019].

\end{frame}

%------------------------------------------------
\begin{frame}

\vspace{3.5cm}
\Huge{\hspace{3.4cm}\usebeamercolor[fg]{title}Thanks!}

\end{frame}

\end{document}

### 2022-02-01 writing/draft.tex
Starting the project

\documentclass[10pt,a4paper,titlepage,twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usepackage{amsmath, amssymb, amsfonts, amsthm, mathtools}
% mathtools for: Aboxed (put box on last equation in align envirenment)
\usepackage{microtype} %improves the spacing between words and letters

\usepackage{graphicx}
%\graphicspath{ {./pics/} {./eps/}}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{ulem}
\usepackage{outlines}

\newcommand\given[1][]{\:#1\vert\:}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% COLOR DEFINITIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[svgnames]{xcolor} % Enabling mixing colors and color's call by 'svgnames'
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\definecolor{MyColor1}{rgb}{0.2,0.4,0.6} %mix personal color
\definecolor{newred}{HTML}{C5000B} %mix personal color

\newcommand{\textb}{\color{Black} \usefont{T1}{lmss}{m}{n}}
\newcommand{\blue}{\color{MyColor1} \usefont{T1}{lmss}{m}{n}}
\newcommand{\blueb}{\color{MyColor1} \usefont{T1}{lmss}{b}{n}}
\newcommand{\red}{\color{newred} \usefont{T1}{lmss}{m}{n}}
\newcommand{\green}{\color{Turquoise} \usefont{T1}{lmss}{m}{n}}
\newcommand{\bluen}{\color{MyColor1} \usefont{T1}{phv}{b}{n}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[norule,hang,flushmargin]{footmisc}
\usepackage[font=footnotesize, hang]{caption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% FONTS AND COLORS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    SECTIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{titlesec}
\usepackage{sectsty}
%%%%%%%%%%%%%%%%%%%%%%%%
%set section/subsections HEADINGS font and color
\sectionfont{\color{MyColor1}}  % sets colour of sections
\subsectionfont{\color{MyColor1}}  % sets colour of sections

%set section enumerator to arabic number (see footnotes markings alternatives)
\renewcommand\thesection{\arabic{section}.} %define sections numbering
\renewcommand\thesubsection{\thesection\arabic{subsection}} %subsec.num.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CAPTIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{caption}
\usepackage{subcaption}
%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup[figure]{labelfont={color=MyColor1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		!!!EQUATION (ARRAY) --> USING ALIGN INSTEAD
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%using amsmath package to redefine eq. numeration (1.1, 1.2, ...) 
%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\theequation}{\thesection\arabic{equation}}

%set box background to grey in align environment 
\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\makeatletter
\patchcmd{\@Aboxed}{\boxed{#1#2}}{\colorbox{black!15}{$#1#2$}}{}{}%
\patchcmd{\@boxed}{\boxed{#1#2}}{\colorbox{black!15}{$#1#2$}}{}{}%
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand{\rmdefault}{phv} % arial
\renewcommand{\sfdefault}{phv} % arial

% Page layout (geometry)
%\setlength\voffset{-1in}
%\setlength\hoffset{-1in}
%\setlength\topmargin{2.5cm}
%\setlength\rightmargin{1.5cm}
%\setlength\oddsidemargin{2.5cm}
%\setlength\textheight{22.4cm}
%\setlength\textwidth{14.00cm}
%\setlength\footskip{1.3cm}
%\setlength\headheight{0cm}
%\setlength\headsep{0cm}

\usepackage
[
        a4paper,% other options: a3paper, a5paper, etc
        left=2.5cm,
        right=2.5cm,
        top=3cm,
        bottom=3cm,
]
{geometry}

% Paragraph formatin
\linespread{1}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 3pt minus 3pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		Import listings package and define parameters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{listings}
\definecolor{grey}{rgb}{0.9,0.9,0.9}
\definecolor{mygreen}{rgb}{0,0.6,0}
\lstset{backgroundcolor=\color{grey},frame=single, language=Matlab, basicstyle=\tiny,commentstyle=\color{mygreen},keywordstyle=\color{blue}
}

% Yma colors
\definecolor{ymagray}{rgb}{0.2,0.2,0.2}
\definecolor{ymablue}{RGB}{0,132,209}
\definecolor{ymadblue}{RGB}{0,69,134}
\definecolor{ymagreen}{HTML}{579d1c}
\definecolor{ymaorng}{HTML}{FF950E}

\definecolor{ymared}{HTML}{c5000b}

%
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}

% - - - - - - - - - - Fancy Header - - - - - - - - - - %
%% L/C/R denote left/center/right header (or footer) elements
%% E/O denote even/odd pages
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[LO,RE]{\slshape\thepage}
\renewcommand{\headrulewidth}{0.1pt}
\cfoot{}

\makeatletter
\let\reftagform@=\tagform@
\def\tagform@#1{\maketag@@@{(\ignorespaces\textcolor{ymadblue}{#1}\unskip\@@italiccorr)}}
\renewcommand{\eqref}[1]{\textup{\reftagform@{\ref{#1}}}}
\makeatother
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor={ymadblue}}

\newcommand{\tilv}[1]{\textbf{\color{ymagreen} [#1]}}
\newcommand{\note}[1]{\textbf{\color{ymagray}[#1]}}
\newcommand{\highlight}[1]{{\bluen{#1}}}
\newcommand{\hghlght}[1]{\textbf{\color{ymaorng} #1}}
\newcommand{\point}[1]{\textbf{\color{ymared} #1}}
\newcommand{\rem}[1]{{\red{\textbf{REMOVE:}} \red{\sout{#1}}}}

\usepackage[onehalfspacing]{setspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Hyphenation correction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\babelhyphenation[icelandic]{
%  Forsendurnar
%  líkinda
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% PREPARE TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\blue Thesis - draft \\
\blueb NB-IoT}
\author{Sigurgeir Gunnarsson \\KTH - Communication Systems}
\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\thispagestyle{empty}
\tableofcontents

%\cleardoublepage
\newpage
\setcounter{page}{1}

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Inngangur
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Intro}

%\hglght{* Describe the problem at hand}

The idea of having all devices connected has led to the concept of Internet of Things (IoT). Here all devices such as electric or water meters, household devices and about every thing with built in sensors, are communicating with each other or with a central server via short distance communication or wide area networks (WAN).

Making this happening brings out a challenge due to the amount of devices, connected to a single cell in the cellular network. All these devices extract resources, they interfere with each other and their location is very probable deep in the house or in the cellar, where communication is hard to withheld.

\point{The challenge is on the physical channels, when the load increases. The random access can result in collision, making it necessary to resend the package. Re-transmission causes unnecessary communication and extra battery. This is therefore something to avoid.}

\point{Keep in mind that NB-IoT is UL-driven service, unlike traditional mobile services that are DL-driven}

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Kafli
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{NB-IoT}

According to 3GPP there are three implementations of wide area network (WAN) narrowband internet-of-things (NB-IoT). One uses a single GSM transmission channel (TCH) and is integrated into the GSM network. The other two are co-existing with LTE where one uses the guard band between LTE channels and the other is embedded into the LTE carrier, using resources of the LTE carrier.

NB-IoT downlink is defined to use 12 sub-carriers, with each sub-carrier having bandwidth of 15 kHz. The uplink uses either a single sub-carrier, which can be of varying bandwidth: 3,75 kHz or 15 kHz, or it uses 3, 6, or 12 sub-carriers of 15 kHz bandwidth each\cite{3gpp}.

The total bandwidth of the downlink sums to 180 kHz. That fits neatly into a GSM TCH which is 200 kHz wide. The LTE guard band varies in width, depending on carrier bandwidth. Most deployments of LTE use 5 MHz or larger bandwidth, since the smaller bandwidths introduce limitations. The sub-carrier definition for the in-band case, is the same as used by the LTE control channels. The NB-IoT transmission can therefore be carried alongside normal LTE traffic if there is capacity to share the control space \point{Need to confirm if it is correctly understood that LTE and NB-IoT share the control space and if they can co-exist on the same channel}.

NB-IoT is defined based on the LTE carrier structure. The smallest LTE transmission element is resource element (RE) which uses 15 kHz bandwidth. A resource block (RB) is 12 RE and occupies 180 kHz (12*15 kHz=180kHz). The LTE carrier can be of varying bandwidth, from 1,4 MHz to 20 MHz, where the difference lies in the number of RB used.

Since the structure is common, NB-IoT can be defined as functionality within the LTE carrier. According to Schlienz (2106) \cite{schlienz} there are specific RB indexes where NB-IoT can be activated. That way both functionalities (LTE and NB-IoT) are active at the same time.

\subsection{Evaluation points}

The limitations when it comes to serving multiple users lies within the random access procedure. The user has limited power and has therefore to make its transmission based on most efficient transmission methods. Collision is important to avoid. Modulation and coding scheme have to navigate between giving the maximum throughput but with best recovery of distorted bits \point{this should be looked into - find articles evaluating this trade-off}.

When the user has gotten the random access procedure completed and is scheduled on the transmission channels. The next step is the focus of the thesis - how to maximise ??, to be able to service more users and withholding the goal of minimum transmission.w

\subsection{Extras}

Schlienz (2016) \cite{schlienz} is a very good summary of NB-IoT structure. The references are all 3GPP TS36.2xx, which is the channel structure of LTE - with a chapter covering NB-IoT.

Question: is it sufficient to rewrite Shclienz (2016) and refer to the article, plus the 3GGP standard ?

\point{* Describe NB-IoT}

There are three implementation methods of NB-IoT. There is standalone, guardband and in-band.

\textbf{6.4.1 Standalone Deployment:}
Standalone deployment is a deployment scenario in which operators deploy NB-IoT using
existing idle spectrum resources. These resources can be the operator’s spectrum
fragments with non-standard bandwidths or spared from other radio access technologies
(RATs) by refarming.

\textbf{6.4.2 LTE Guardband Deployment:}
Guardband deployment is a deployment scenario in which operators deploy NB-IoT in guard
bands within existing LTE spectrum resources.

\textbf{6.4.3 LTE In-band Deployment:}
In-band deployment is a deployment scenario in which operators deploy NB-IoT using
existing LTE in-band resource blocks (RBs).
\tilv{GSMA deployment guide}

In the 3GPP NB-IoT technical specification, chapter 4.2, the radio resources are defined in deep details (if relevant).
\cite{schlienz}.

In the following work of the focus will be on in-band LTE NB-IoT.

\note{WHERE IS TTI DESCRIBED !!!}

\point{* Cover the access method}

Schlienz (2016) is a really god reference to work the access method from.
1

* Coverage classes - how different classes affect the access mechanism

Here are some outstanding questions. One coverage class something mentioned in the specifikation or is it something suggested in literature.

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Heimildaskrá
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{References}

\begin{thebibliography}{4}
\bibitem{schlienz}
Schlienz, J., \& Raddino, D. (2016).
\textit{Narrowband internet of things whitepaper}
. White Paper, Rohde\&Schwarz, 1-42.

\bibitem{3gpp}
3GPP TS 36.201,
\textit{3rd Generation Partnership Project; Technical Specification Group Radio Access Network; Evolved Universal Terrestrial Radio Access (E-UTRA); LTE physical layer; General description}, version 15.2.0, Release 15.

\bibitem{3gpp211}
3GPP TS 36.211,
\textit{3rd Generation Partnership Project; Technical Specification Group Radio Access Network; Evolved Universal Terrestrial Radio Access (E-UTRA); LTE physical layer; General description}, version 15.14.0, Release 15.

\bibitem{chen}
Chen, M., Miao, Y., Hao, Y., \& Hwang, K. (2017).
\textit{Narrow band internet of things}
. IEEE access, 5, 20557-20577.

\end{thebibliography}

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Glósur !!!!
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{\point{Outstanding}}

\begin{itemize}
\item Is there a code available that makes it possible to emulate the access mechanism and show the clash when the load increases ?

\end{itemize}

\end{document}

### 2022-02-01 writing/nb-iot.tex
Starting the project

\documentclass[10pt,a4paper,titlepage,twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usepackage{amsmath, amssymb, amsfonts, amsthm, mathtools}
% mathtools for: Aboxed (put box on last equation in align envirenment)
\usepackage{microtype} %improves the spacing between words and letters

\usepackage{graphicx}
%\graphicspath{ {./pics/} {./eps/}}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{ulem}
\usepackage{outlines}

\newcommand\given[1][]{\:#1\vert\:}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% COLOR DEFINITIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[svgnames]{xcolor} % Enabling mixing colors and color's call by 'svgnames'
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\definecolor{MyColor1}{rgb}{0.2,0.4,0.6} %mix personal color
\definecolor{newred}{HTML}{C5000B} %mix personal color
\newcommand{\textb}{\color{Black} \usefont{T1}{lmss}{m}{n}}
\newcommand{\blue}{\color{MyColor1} \usefont{T1}{lmss}{m}{n}}
\newcommand{\blueb}{\color{MyColor1} \usefont{T1}{lmss}{b}{n}}
\newcommand{\red}{\color{newred} \usefont{T1}{lmss}{m}{n}}
\newcommand{\green}{\color{Turquoise} \usefont{T1}{lmss}{m}{n}}
\newcommand{\bluen}{\color{MyColor1} \usefont{T1}{phv}{b}{n}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[norule,hang,flushmargin]{footmisc}
\usepackage[font=footnotesize, hang]{caption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% FONTS AND COLORS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    SECTIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{titlesec}
\usepackage{sectsty}
%%%%%%%%%%%%%%%%%%%%%%%%
%set section/subsections HEADINGS font and color
\sectionfont{\color{MyColor1}}  % sets colour of sections
\subsectionfont{\color{MyColor1}}  % sets colour of sections

%set section enumerator to arabic number (see footnotes markings alternatives)
\renewcommand\thesection{\arabic{section}.} %define sections numbering
\renewcommand\thesubsection{\thesection\arabic{subsection}} %subsec.num.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CAPTIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{caption}
\usepackage{subcaption}
%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup[figure]{labelfont={color=MyColor1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		!!!EQUATION (ARRAY) --> USING ALIGN INSTEAD
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%using amsmath package to redefine eq. numeration (1.1, 1.2, ...) 
%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\theequation}{\thesection\arabic{equation}}

%set box background to grey in align environment 
\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\makeatletter
\patchcmd{\@Aboxed}{\boxed{#1#2}}{\colorbox{black!15}{$#1#2$}}{}{}%
\patchcmd{\@boxed}{\boxed{#1#2}}{\colorbox{black!15}{$#1#2$}}{}{}%
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand{\rmdefault}{phv} % arial
\renewcommand{\sfdefault}{phv} % arial

% Page layout (geometry)
%\setlength\voffset{-1in}
%\setlength\hoffset{-1in}
%\setlength\topmargin{2.5cm}
%\setlength\rightmargin{1.5cm}
%\setlength\oddsidemargin{2.5cm}
%\setlength\textheight{22.4cm}
%\setlength\textwidth{14.00cm}
%\setlength\footskip{1.3cm}
%\setlength\headheight{0cm}
%\setlength\headsep{0cm}

\usepackage
[
        a4paper,% other options: a3paper, a5paper, etc
        left=2.5cm,
        right=2.5cm,
        top=3cm,
        bottom=3cm,
]
{geometry}

% Paragraph formatin
\linespread{1}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 3pt minus 3pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		Import listings package and define parameters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{listings}
\definecolor{grey}{rgb}{0.9,0.9,0.9}
\definecolor{mygreen}{rgb}{0,0.6,0}
\lstset{backgroundcolor=\color{grey},frame=single, language=Matlab, basicstyle=\tiny,commentstyle=\color{mygreen},keywordstyle=\color{blue}
}

% Yma colors
\definecolor{ymagray}{rgb}{0.2,0.2,0.2}
\definecolor{ymablue}{RGB}{0,132,209}
\definecolor{ymadblue}{RGB}{0,69,134}
\definecolor{ymagreen}{RGB}{87,157,28}

%
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}

% - - - - - - - - - - Fancy Header - - - - - - - - - - %
%% L/C/R denote left/center/right header (or footer) elements
%% E/O denote even/odd pages
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[LO,RE]{\slshape\thepage}
\renewcommand{\headrulewidth}{0.1pt}
\cfoot{}

\makeatletter
\let\reftagform@=\tagform@
\def\tagform@#1{\maketag@@@{(\ignorespaces\textcolor{ymadblue}{#1}\unskip\@@italiccorr)}}
\renewcommand{\eqref}[1]{\textup{\reftagform@{\ref{#1}}}}
\makeatother
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor={ymadblue}}

\newcommand{\note}[1]{\textcolor{ymagray}{\textbf{[#1]}}}
\newcommand{\highlight}[1]{{\bluen{#1}}}
\newcommand{\hglt}[1]{{\red{#1}}}
\newcommand{\rem}[1]{{\red{\textbf{REMOVE:}} \red{\sout{#1}}}}

\usepackage[onehalfspacing]{setspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Hyphenation correction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\babelhyphenation[icelandic]{
%  Forsendurnar
%  líkinda
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% PREPARE TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\blue Thesis - draft \\
\blueb NB-IoT}
\author{Sigurgeir Gunnarsson \\KTH - Communication Systems}
\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\thispagestyle{empty}
\tableofcontents

\cleardoublepage
\newpage
\setcounter{page}{1}

\section{Introduction}

Problem statement and scope of the work !!!

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Kafli
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{NB-IoT}

* Describe NB-IoT

* Cover the access method

* Coverage classes - how different classes affect the access mechanism

* Cover power usage of the access procedure and why it matters

* Describe the problem at hand

Is there a code available that makes it possible to emulate the access mechanism and show the clash when the load increases ?

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Kafli
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{Machine Learning}

* Machine Learning (ML) mechanism to be investigated and tried out

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Kafli 2
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\if(0)
\clearpage
\section{Beiting hefðbundnu aðferðarinnar}\label{gr_main}

Í kafla \ref{gr_yfirf} er farið yfir grunninn bak við aðferð Gutenberg-Richter (GR) og hvernig á að reikna b-gildið, óþekktu stærðina í jöfnu líkindafallsins. Til að fá tölulegt mat á aðferðina er skoðað safn jarðskjálfta á Reykjanesi og Reykjaneshrygg. Safnið inniheldur alla jarðskjálfta á árunum 1991 til 2015 með minnsta marktæka jarðskjálftann M$_{min}=2$. Mynd \ref{fig:collect} sýnir hvernig jarðskjálftar í safninu af stærð M$\geq3,7$ dreifast á kort, litað eftir stærð jarðskjálfta.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{img/dreifing_kort.png}
  \vspace{-4mm}
  \caption{Dreifing jarðskjálfta lagt út á kort, minnsta stærð jarðskjálfta er M$=3,7$}
  \label{fig:collect}
\end{figure}

Þar sem GR lögmálið lýsir sambandinu milli stærðar jarskjálfta og fjölda jarðskjálfta af tiltekinni stærð, er eðlilegt að skoða dreifinguna. Mynd \ref{fig:gr_dist} sýnir annars vegar fjölda jarðskjálfta af tiltekinni stærð á línulegum skala (efra grafið) og hins vegar á log-normal skala (neðra grafið).

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{img/gr_dist.png}
  \vspace{-4mm}
  \caption{Samband stærðar jarðskjálfta og fjölda af tiltekinni stærð}
  \label{fig:gr_dist}
\end{figure}

Ólíkt GR lögmálinu, þá sýnir mynd \ref{fig:gr_dist} ekki uppsafnaðan fjölda. Það má engu að síður sjá veldishnignun á efra grafinu og beina línu á vinstri hluta neðra grafsins. Beina línan virðist hins vegar einungis eiga við M upp í kringum $3,5$. Til samanburðar er mynd \ref{fig:cum_sum} sem sýnir sambandið eins og því er lýst með lögmáli GR, þ.e. með uppsöfnuðum fjölda af sömu stærð jarðskjálfta eða stærri.

%\begin{figure}[h]
%  \centering
%  \includegraphics[width=\textwidth]{img/gr_log_norm_dist.png}
%  \vspace{-4mm}
%  \caption{Dreifingin sett fram skv. lögmáli Gutenberg-Richter}
%  \label{fig:gr_log_norm_dist}
%\end{figure}

Mynd \ref{fig:cum_sum} sýnir mun skýrara línulegt samband en mynd \ref{fig:gr_dist}, enda mælingar af tiltekinni stærð orðnar tengdar mælingum af stærðunum fyrir ofan. Þetta hefur einhvers konar síunar áhrif þar sem vinstri hlutinn á neðra grafi myndar \ref{fig:gr_dist} er orðinn mun hreinni á mynd \ref{fig:cum_sum}.

Næsta skref er að beita aðferð kafla \ref{gr_yfirf} til að fá b-gildið eða hallatöluna á mynd \ref{fig:cum_sum}. Hallatalan lýsir línulega sambandinu milli stærðar og uppsafnaðs fjölda og fæst með jöfnu \ref{eq:mle}: $$b=0.93512$$ Til að fá skurðpunkt við y-ás er notuð línuleg aðhverfugreining með gefið b. Línulega sambandið er sýnt með brotalínu á mynd \ref{fig:cum_sum}.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{img/cum_sum.png}
  \vspace{-4mm}
  \caption{Dreifingin ásamt línulega sambandi GR, b=0.93512}
  \label{fig:cum_sum}
\end{figure}

Til að fá samhengi í b-gildið eru aftur reiknuð líkindin, nú fyrir jarðskjálfta af stærð $M\geq5$ og $M\geq6$: $$P[5\leq m]=10^{-0.93512\cdot(5-2)}=0.001565=0.1565\%$$ $$P[6\leq m]=10^{-0.93512\cdot(6-2)}=0.0001818=0.01818\%$$ Niðurstaðan er að líkindin hafa aukist fyrir $M\geq6$. Í stað eins af hverjum 10000 jarðskjálftum sé $M\geq6$, þá er búist við að einn af hverjum 5501 sé það. Þetta er í samræmi við líkindadreifinguna en dreifingin fellur minna þegar b-gildið lækkar og þar með aukast líkindin í halanum þ.e. fyrir stærri M.
%Áætlaður endurkomutími jarðskjálfta af stærð M=6 eða stærri, út frá jöfnu \ref{eq:prob_ex}, er reiknaður 15.1 ár.

Gagnasafnið er 7271 mælingar og spannar 9015 daga. Það gerir að meðaltali 0,8065 jarðskjálfta á dag. Líkindin gefa að 1 af hverjum 639 jarðskjálftum uppfylli $M\geq5$, sem jafngildir endurkomutíma upp á 17 mánuði. Á sama hátt, miðað við að 1 af hverjum 5501 uppfylli skilyrðin $M\geq6$, er endurkomutími jarðskjálfta $M\geq6$ rúm 12 ár.

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Kafli 3
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{Aðferð stórra útgilda - Pareto dreifing}

Í bókinni \textit{An Introduction to Statistical Modeling of Extreme Values} eftir Coles, S. (2001) er fjallað um vandamálið sem kemur upp á myndum \ref{fig:gr_dist} og \ref{fig:cum_sum}. Þá er verið að vísa til þess hvernig matið á dreifingunni bjagast og beygir af leið þegar fjöldi tilfella verður lítill fyrir gildi langt frá massanum.

%Hér verður unnið með dreififall F og þröskuld u þannig að líkindin á atburði sem fer umfram þröskuldinn er: $$P[X>u+y \given[\big] X>u] = \frac{1-F(u+y)}{1-F(u)}, \quad y>0$$

Lausnin sem Coles kynnir, byggir á dreifingum sem mynda fjölskyldu almennra útgilda dreifinga (e. generalized extreme value distribution). Í tilfelli þar sem $X_1,X_2,...$ eru óháðar hendingar með sameiginlega dreifingu $F$, þá er hægt að skilgreina $M_n$ þannig:
\[
M_n = \max\{X_1,...,X_n\}
\]
Ef líkindin uppfylla $P[M_n<z]\approx G(z)$, þá er hægt að setja fram almennu formúlana fyrir dreifinguna $G(z)$:
\[
G(z) = \exp \left\lbrace -\left[ 1+ \xi \left( \frac{z-\mu}{\sigma} \right) \right]^{-1/\xi} \right\rbrace
\]
þar sem z er stak í menginu $\{z : 1 + \xi (z-\mu)/ \sigma > 0\}$ og breiturnar uppfylla $-\infty < \mu,\xi < \infty$ og $0<\sigma$. Sértilfelli af formúlunum, sem á við verkefnið sem er verið að skoða, er þegar $0<\mu,\sigma$. Þá er hægt að finna nógu stórt $u$ þannig að dreifinguna $(X-u)$, gefið $X>u$, er hægt að nálga með:
\begin{equation}\label{eq:gen_par}
H(y) = 1 - \left( 1 + \frac{\xi y}{\tilde{\sigma}} \right) ^{-1/\xi}
\end{equation}
þar sem y er stak í menginu $\{y : y>0 ~og~ (1 + \xi y/ \tilde{\sigma}) > 0\}$ og
\begin{equation}\label{eq:sigma}
\tilde{\sigma} = \sigma + \xi(u-\mu)
\end{equation}
%\note{S.C.[2001] kafli 4.2}

Jafna \ref{eq:gen_par} gengur undir heitinu almenn Pareto dreifing (Generalized Pareto family). Sértilfelli hennar er þegar $\xi\rightarrow0$ en þá fæst:
\[
H(y) = 1 - \exp\left(-\frac{y}{\tilde{\sigma}} \right),~ y>0
\]
Hér er komin formúla fyrir veldisdreifingu með stuðul $1/\tilde{\sigma}$, $Exp(1/\tilde{\sigma})$.

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Kafli 6
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{Samanburður við gagnasafn sem nær lengra í tíma}

Veikleiki niðurstaðnanna um tíðni jarðskjálfta af stærð $M\geq6$ liggur í því að verið er að skoða stærð jarðskjálfta sem ekki fyrirfinnst í gagnasafninu. Til að bæta upp fyrir veikleikann gagnasafnsins er líka skoðaður endurkomutími jarðskjálfta af stærð $M\geq5$.

Stærðin $M=6$ er hins vegar ákveðin út frá markverðanleika. Jarðskjálftar minni en 6 teljast hóflegir skv. USGS og hafa litla verkfræðilega þýðingu. Stærri jarðskjálftar en 6 teljarst sterkir og fara að gera sig gildandi þegar kemur að hönnun mannvirkja.

Gagnasafnið er mjög ýtarlegt. Það hefur upplausn $\Delta M=0,01$ og minnsta marktæka jarðskjálfta $M_{min}=2$. Hins vegar spannar það ekki nema tæp 25 ár, frá 1991 til 2015. Til að fá stærra tímabil þarf að skoða gagnasafn sem spannar rúm 64 ár, frá 1926 - 1990. Munurinn er að upplausnin á lengra safninu er ekki nema $\Delta M=0,1$ og minnsti marktæki jarðskjálftinn $M_{min}=3$. Í lengra safninu eru hins vegar þrír jarðskjálftar af stærðinni $M\geq6$, á sama svæði og ýtarlegra gagnasafnið þekur.

Enn annað gagnasafn byggir á mati út frá sögulegum heimildum og er gefið í töflu í Náttúruvá á Íslandi (Júlíus Sólnes o.fl. (2013)). Þetta safn nær frá síðasta stóra jarðskjálfta 2008 aftur til 1706 og inniheldur 21 jarðskjálfta stærri eða jafnt og 5,8. Á svæðinu sem ýtarlega safnið þekur eru sjö jarðskjálftar $M\geq 6$ á árunum frá 1706 til 1968.

\begin{table}[h]
  \centering
  \begin{tabular}{|l|p{1.4cm}p{1.4cm}p{1.4cm}p{1.4cm}p{1.4cm}|p{1.2cm}p{1.2cm}p{1.2cm}|}
  \hline
        & \multicolumn{5}{|c|}{Dreififöll} & \multicolumn{3}{|c|}{Gagnasöfn} \\
        & GR & GR & GR & Pareto & Pareto & 1991 -& 1926 - & 1706 -\\
     Stærð & $M=2,0$ & $M=4,0$ & $M=4,4$ & $M=3$ & $M=4,2$ & 2015 & 1990 & 2008\\
  \hline
  $M\geq5$ & $1,41$ & $3,62$ & $11,6$ & $4,26$ & $13,16$ & $3,53$ & $1,34$ & - \\
  $M\geq6$ & $ 12,16$ & $42,75$ & $203,09$ & $204,02$ & $296,07$ & - & $21,39$ & $43,19$ \\
  \hline
  \end{tabular}
  \caption{Samantekt útreiknaðs endurkomutíma skv. dreififöllum og endurkomutími þriggja gagnasafna}
  \label{table:prob}
\end{table}

Útreiknaður endurkomutími jarðskjálfta af stærð $M\geq5$ og $M\geq6$ sem hefur verið reiknaður í fyrri köflum er tekinn saman í töflu \ref{table:prob}. Til samanburðar er endurkomutíminn eins og hann kemur fyrir í gagnasöfnunum þremur. Strikin í gagnasöfnunum koma til af því að í 1991-2015 safninu er enginn jarðskjálfti $M\geq6$. Í 1706-2008 safninu er minnsti marktæki jarðskjálftinn 5,8 þannig safnið er ekki marktækt fyrir $M\geq5$.

Markverðasta niðurstaðan úr töflu \ref{table:prob} er að GR - $M=4,0$ kemst næst endurkomutímunum í gagnasöfnunum. GR - $M=2,0$ passar vel við 64 ára safnið en er langt frá báðum hinum. Pareto - $M=3,0$ gefur svipaða niðurstöðu og GR - $M=4,0$ á endurkomutíma jarðskjálfta $M\geq5$ en er langt frá endurkomutímanum á jarðskjálfta $M\geq6$. Hinar aðferðirnar voru hvergi nálægt því að gefa nothæfa niðurstöðu.

Í kafla \ref{comp} var dreififallið fyrir GR - $M=2$ afskrifað þar sem það vék frá mælingunum þegar $M$ stækkaði. Það skýtur því skökku við að líkindin sem dreififallið falli svo vel að endurkomutíma 1926-1990 safnsins. Til að sjá hvernig þetta gerist þarf að skoða dreifingu mælinga úr 1926-1990 safninu saman með 1991-2015 safninu á einni mynd.

Mynd \ref{fig:gagnasofn} sýnir hvernig lengra gagnasafnið vex mun hraðar fyrir stór gildi $M$. Það skýrir einnig styttri endurkomutíma safnsins samanborið við hin, þ.e. það er meiri fjöldi mælinga af þessari stærð í safninu.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{img/gagnasofn.png}
  \vspace{-4mm}
  \caption{Dreifing mælinga fyrir gagnasöfnin 1926-1990 og 1991-2015}
  \label{fig:gagnasofn}
\end{figure}

Hægt hefði verið að reikna b-gildi aðferðar Gutenberg-Richter með því að nota bæði gagnasöfnin. Niðurstaðan hefði líklegast verið í áttina að því sem fékkst með ótakmarkaða gagnasafninu. Hins vegar hefði þurft að gera sérstakar ráðstafanir þar sem upplausnin og tímaspan safnanna er ólík og fjöldinn þess vegna ekki samanburðarhæfur beint.

Í grein Weichert (1980) flokkar hann jarðskjálftana í stærðarflokka, reiknar líkindi flokksins og fær b-gildið út úr margfeldi líkindanna deilt með margfeldi fjöldans í hverjum flokki. Þessi aðferð er mikilvæg þar sem gagnasöfnin eru mjög misjöfn. Tímalengdin er ólík, upplausnin oft minni og minnsti marktæki jarðskjálftinn stærri í eldri gagnasöfnum.

Það sem mynd \ref{fig:gagnasofn} sýnir einnig er hvernig virknin er ólík milli mismunandi tímabila. Fjöldinn þegar $M\geq5,5$ er t.d. miklu meira en þrisvar sinnum meiri í eldra gagnasafninu jafnvel þó safnið sé ekki nema þrisvar sinnum lengra í tíma. Einnig jarðskjálftarnir $M\geq6$ sem eru ekki til staðar í stutta gagnasafninu, þrátt fyrir að endurkomutíminn ætti að vera 21 ár skv. töflu \ref{table:prob}.

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Kafli 7
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{Afþyrping}

Jarðskjálftum er almennt skipt í fjóra flokka: fyrirskjálfta (e. foreshock), aðalskjálfta (e. mainshock), eftirskjálfta (e. aftershock) og sveima (e. swarm). Fyrstu þrír flokkarnir tengjast þar sem fyrir- og eftirskjálftar gerast samfara einhverjum atburði sem er þá aðalskjálftinn. Sveimarnir eru órói svipaður eftirskjálftum en tengjast engum aðalskjálfta.

Fyrir utan Gutenberg-Ricther, þá eru tvær kenningar til viðbótar sem tengja flokkana þrjá: lögmál Båth og lögmál Omori. Lögmál Båth fullyrðir að það sé fastur stærðarmunur á aðalskjálftanum og fyrsta stóra eftirskjálftanum \note{Má vísa í grein Båth ?}. Lögmál Omori er dreififall sem segir til um hversu mikið eftirskjálftarnir minnka í stærð eftir því sem tíminn líður \note{Má vísa í grein Omori þar sem hann setur þetta lögmál fram ? Omori, F. (1894) On the Aftershocks of Earthquakes. Journal of the College of Science, Imperial University of Tokyo, 7, 111-120.}.

Með þessum lögmálum er hægt að einangra aðalskjálftann sem er aðal orsök atburðarins og jafnframt stærsti jarðskjálftinn. Knopoff (2000) bendir einnig á að við hermun jarðskjálftasafna er verið að búa til jarðskjálfta sem allir hafa eðlisfræðilega eiginleika aðalskjálfta. Það er langt frá veruleikanum enda fylgir umtalsverður fjöldi eftirskjálfta hverjum aðalskjálfta. Þess vegna leggur hann mikla áherslu á að einangra safnið við aðalskjálfta með afþyrpingu (e. declustering).

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{img/cluster.png}
  \vspace{-4mm}
  \caption{Jarðskjálftahrina í Ölfusi, nóvember 1998}
  \label{fig:cluster}
\end{figure}

Mynd \ref{fig:cluster} sýnir atburð sem átti sér stað í nóvember 1998 í Ölfusi. Það er augljóslega einn aðalskjálfti og fjórir sem ekki er augljóst hvort séu aðalskjálftar eða eftirskjálftar. Hinir 106 skjálftarnir eru svo eftirskjálftar sem Knopoff myndi fjarlægja fyrir sína greiningu.

Afþyrping myndi hafa umtalsverð áhrif á þá greiningu sem hefur verið gerð í fyrri köflum. Fjöldi jarðskjálfta $M<5$ myndi minnka umtalsvert og safnið myndi ekki ná eins langt niður í stærð. B-gildið sem fengist með slíku safni væri örugglega minna en það sem fékkst í kafla \ref{gr_main} þar sem brattinn á fjöldanum í safninu væri minni. Hallinn á mælingunum $M\geq5$ væri hins vegar sá sami.

Áhrifin af afþyrpingunni er að vankantarnir sem hafa verið nefndir við beinu línu GR dreifingarinnar yrðu ýktari. Bein lína nær aldrei að falla bæði að litlu gildunum og stóru gildunum sem hafa annan halla. Auk þess er hætt við að sambandið milli stærðar og fjölda af ákveðinni stærð verði ekki lengur línulegt.

Þetta eru hins vegar meðmæli með Pareto dreifingunni sem skv. mynd \ref{fig:sigma_xi} hefur sterka tilhneigingu til að vera hvelft ($\xi$ gildin flest fyrir neðan x-ás). Ef línulega sambandið hverfur er líklegast að samband stærðar og fjölda yrði hvelft eins og Pareto dreifingin.

%\note{Með því að beita afþyrpingu (e. declustering) fækkar mælingunum fyrir neðan stærstu gildin ,,Båth's-law''. Þá myndi ég ætla að hallinn minnki því fjöldinn vex ekki eins hratt. Það eru rök með Pareto dreifingunni því hún er hvelft og getur þess vegna fallið að minni mælingunum ásamt stærri mælingunum.}

%\note{Hugleiðingar um vankant Pareto þegar hún fer inn á ókannaða svæðið - stærstu M gildin. Hvelfti eiginleikinn gerir það að verkum að dreifingin er svartsýnni fyrir stór M en hinar aðferðirnar. Sér í lagi gildir það fyrir GR þegar hallatalan á safninu minnkar. Þá er jafnvel spurning hvort GR verði ekki of bjartsýn.}

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Niðurstöður
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{Niðurstöður}

Það virðist sem aðferð Gutenberg-Richter skili viðunandi niðurstöðu þegar meta á jarðskjálfta af ákveðinni stærð. Í tilfelli mats á jarðskjálftum af stórum stærðum, þar sem fjöldinn í safninu er takmarkaður, þarf að gera ráðstafanir eins og að takmarka gagnasafnið að neðan í stærð jarðskjálfta sem eru notaðir við mat stika í GR módelinu.

Hugmynd verkefnisins var að skoða Pareto dreifingu en hún er viðurkennd dreifing þegar safnið fylgir veldisdreifingu og skoða á líkindin á útgildum.

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Heimildaskrá
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{Heimildaskrá}

Aki, K. (1965). Maximum likelihood estimate of b in the formula $\log~N=a-bM$ and its confidence limits. \textit{Bulletin of the earthquake research institute}, 43, 327-239

Coles, S. (2001). \textit{An introduction to statistical modeling of extreme values} London: Springer.

Gutenberg, B. \& Richter, C. (1954). \textit{Seismicity of the Earth and Associated Phenomena, 2. útgáfa}. Princeton, New Jersey: Princeton University Press

Hyndman, R. J. \& Koehler, A. B. (2006). Another look at measures of forecast accuracy. \textit{International Journal of Forecasting}, 22(4), 679-688

Júlíus Sólnes, Freysteinn Sigmundsson og Bjarni Bessason (2013). \textit{Náttúruvá á Íslandi}. Viðlagatrygging Íslands / Háskólaútgáfan

Knopoff, L. (2000). The magnitude distribution of declustered earthquakes in Southern California. \textit{Proceedings of the National Academy of Sciences}, 97(22), 11880-11884

Marzocchi, W. \& Sandri, L. (2003). A review and new insights on the estimation of the b-value and its uncertainty. \textit{Annals of geophysics}, 46(6), 1271-1282

Marzocchi, W. \& Sandri, L. (2007). A technical note on the bias in the estimation of the b-value and its uncertainty through the Least Squares technique. \textit{Annals of geophysics}, 50(3), 329-339

Rasch, D. (1977). Sample size determination for estimating the parameter of an exponential distribution. \textit{Biometrical Journal}, 19(7), 521-528

Utsu, T. (1965). A method for determining the value of b in a formula $\log~n=a-bM$ showing the magnitude-frequency relation. \textit{Geophysics bulletin of the Hokkaido University}, 13, 99-103

Weichert, D. (1980) Estimation of the earthquake recurrence parameters for unequal observation periods for different magnitudes. \textit{Bulletin of the seismological society of America}, 70(4), 1337-1346

% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
% Appendix
% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\clearpage
\section{Viðauki}

\subsection{Fjöldi í flokki}\label{fj_flokki}

Við úrvinnslu jarðskjálftamælinga og við mat á b, er oftar en ekki notast við flokkaðar mælingar. Þá er átt við að M sé skipt í flokka af ákveðinni stærð og talin fjöldi jarðskjálfta í hverjum flokk. Þessi framsetning er í anda massafalls sem gefur hlutfall gilda í safni af ákveðinni stærð.

Vandinn við þessa aðferð er að meðaltalið, reiknað út frá flokkunum, verður bjagað miðað við meðaltalið út úr gagnasafninu. Einnig er M$_{thres}$ í jöfnu \ref{eq:ml_b} miðgildi minnsta flokksins í gagnasafninu. Þarna er komin viðbótar skekkja sem er reynt að leiðrétta fyrir með því að segja M$_{low}$= M$_{thres}-\Delta$M$/2$ (M\&S[2003]).

Við lausn jöfnu \ref{eq:ml_b} er notast við miðgildi flokkanna M$_0$,...,M$_k$, fjöldann í hverjum flokk n$_0$,..,n$_k$ og leiðréttinguna M$_{low}$= M$_{thres}-\Delta$M$/2$. Jafnan verður þá:
\begin{eqnarray}
b & = & \frac{1}{\ln(10)(\hat{\mu} - M_{thres}+\Delta M/2)}\nonumber\\
  & = & \frac{1}{\ln(10)\left(\frac{1}{N}\sum_{i=0}^k n_iM_i - M_{thres} + \Delta M/2\right)}\nonumber\\
  & = & \frac{1}{\ln(10)\left(\frac{1}{N}\sum_{i=0}^k n_iM_i - M_{low}\right)}
\end{eqnarray}
en hér er N heildarfjöldi jarðskjálfta í gagnasafninu og $\Delta$M stærð flokkanna. Í tilfelli þess að gagnasafnið hafi að geyma minnsta marktæka gildið M$_{min}$, þá má segja að M$_{low}=$ M$_{min}$.

\note{Hér má bæta við texta sem vísar í grein Weichert um að blanda saman söfnum af mismunandi stærð og tíma}

\fi

\end{document}

### 2022-02-03 writing/draft.tex
Starting on the draft

@@
 
-Making this happening brings out a challenge due to the amount of devices, connected to a single cell in the cellular network.
-All these devices extract resources, they interfere with each other and their location is very probable deep in the house or in the cellar, where communication is hard to withheld.
+One implementation is utilising the LTE standard and modifying it to meet the requirement needed for IoT.
+Besides the massive volumes it has to handle, there are requirements on coverage, communication complexity and battery life \cite{wang}.
+
+The devices need to be accessible from hard radio conditions like from house cellar or a technique room in the core.
+The communications have to be simple for the sake of resilience and low power consumption.
+Lastly it is expected that the battery to last the live of the device.
+It is therefore extremely important that no unnecessary transmission is undertaken.
+
+When determining the capacity of the base station, or the serving cell, it is not enough to have the theoretical capacity analysed.
+The random nature of when the device transmits, increases the chance of collision when the number of devices increases.
+As mentioned the requirement is to avoid such collision.
+The capacity therefore reduces when taken this into account.
+
+A suitable approach is to identify the theoretical capacity of an NB-IoT cell.
+Then introducing randomness into the access, taking into account the possibility of collision.
+The result is a multi-dimensional space of capacity weighted against probability of collision and number of devices.
+To further enrich the analysis, the radio condition can be taken into account, introducing further randomness since poorer radio condition results in higher re-transmission probability when the number of devices increases.
+
+% Making this happening brings out a challenge due to the amount of devices, connected to a single cell in the cellular network.
+All these devices extract resources, they interfere with each other and their location is probably deep in the house or in the cellar, where communication is hard to withheld.
+
+%The standard has taken into account the attenuation in the radio, adding robustness in the transmission providing up to 20 dB higher pathloss.
+When this is combined with the need for long battery life, the reliability of the transmission becomes important.
+If the battery limit is regarded as a finite pool of transmissions, it becomes unnecessary usage to re-transmit.
+
+\vspace{2cm}
 
@@
 One uses a single GSM transmission channel (TCH) and is integrated into the GSM network.
-The other two are co-existing with LTE where one uses the guard band between LTE channels and the other is embedded into the LTE carrier, using resources of the LTE carrier.
+The other two are co-existing with LTE where one uses the guard band between LTE channels and the other is embedded into the LTE channel.
 
-NB-IoT downlink is defined to use 12 sub-carriers, with each sub-carrier having bandwidth of 15 kHz.
+Regardless of where the carrier resides, the physical channel structure is always the same.
+It is based on LTE, but scaled down to accommodate for the reduced need for high transmission throughput.
+That means the carrier bandwidth is narrower, but preserving the ratio between signali resource blocks and transmission blocks.
+
+NB-IoT downlink is defined to use 12 sub-carriers, where each sub-carrier has bandwidth of 15 kHz.
 The uplink uses either a single sub-carrier, which can be of varying bandwidth: 3,75 kHz or 15 kHz, or it uses 3, 6, or 12 sub-carriers of 15 kHz bandwidth each\cite{3gpp}.
@@
 The total bandwidth of the downlink sums to 180 kHz.
-That fits neatly into a GSM TCH which is 200 kHz wide.
+That fits neatly into a GSM transmission channel (TCH) which is 200 kHz wide.
 The LTE guard band varies in width, depending on carrier bandwidth.
-Most deployments of LTE use 5 MHz or larger bandwidth, since the smaller bandwidths introduce limitations.
-The sub-carrier definition for the in-band case, is the same as used by the LTE control channels.
+Most deployments of LTE though, use 5 MHz or larger bandwidth, since the smaller bandwidths introduce exceptions from the normal structure.
+
+%The sub-carrier definition for the in-band case, is the same as used by the LTE control channels.
 The NB-IoT transmission can therefore be carried alongside normal LTE traffic if there is capacity to share the control space \point{Need to confirm if it is correctly understood that LTE and NB-IoT share the control space and if they can co-exist on the same channel}.
 
-NB-IoT is defined based on the LTE carrier structure.
+%NB-IoT is defined based on the LTE carrier structure.
 The smallest LTE transmission element is resource element (RE) which uses 15 kHz bandwidth.
@@
 
-Since the structure is common, NB-IoT can be defined as functionality within the LTE carrier.
+%Since the structure is common, NB-IoT can be defined as functionality within the LTE carrier.
 According to Schlienz (2106) \cite{schlienz} there are specific RB indexes where NB-IoT can be activated.
@@
 \begin{thebibliography}{4}
+\bibitem{wang}
+Wang, Y. P. E., Lin, X., Adhikary, A., Grovlen, A., Sui, Y., Blankenship, Y., ... \& Razaghi, H. S. (2017).
+\textit{A primer on 3GPP narrowband Internet of Things}. IEEE communications magazine, 55(3), 117-123.
+
 \bibitem{schlienz}
 Schlienz, J., \& Raddino, D. (2016).
-\textit{Narrowband internet of things whitepaper}
-. White Paper, Rohde\&Schwarz, 1-42.
+\textit{Narrowband internet of things whitepaper}. White Paper, Rohde\&Schwarz, 1-42.
 
@@
 
+\hghlght{Statement:} For a device with 164 dB coupling loss, a 10-year battery life can be reached if the UE transmits 200-byte data a day on average {[TR 45.820 v13.1.0, “Cellular system support for ultra low complexity and low throughput internet of things,” Nov.
+2015]}
+
 \end{document}

### 2022-02-10 writing/draft.tex
Small update

@@
 Regardless of where the carrier resides, the physical channel structure is always the same.
-It is based on LTE, but scaled down to accommodate for the reduced need for high transmission throughput.
-That means the carrier bandwidth is narrower, but preserving the ratio between signali resource blocks and transmission blocks.
+It is based on LTE, but scaled down to occupy less bandwidth which gives an opportunity to place the carrier with greater flexibilty.
+The need for high throughput isn't relevant for NB-IoT, since the devices send limited amount of data.
+The LTE structure in itself incorporates high throughput.
+This can be utilised by NB-IoT, adding further data protection and reconstruction.
+That gives added capability to handle channel path loss of up to 20 dB {[\hghlght{missing reference}}].
 
-NB-IoT downlink is defined to use 12 sub-carriers, where each sub-carrier has bandwidth of 15 kHz.
+NB-IoT downlink is defined to use 12 sub-carriers, with each sub-carrier having bandwidth of 15 kHz.
+That adds up to 180 kHz, which matches perfectly the GSM transmission channel (TCH) of 200 kHz.
 The uplink uses either a single sub-carrier, which can be of varying bandwidth: 3,75 kHz or 15 kHz, or it uses 3, 6, or 12 sub-carriers of 15 kHz bandwidth each\cite{3gpp}.
+The last is the same as for the downlink and the total bandwidth becomes 180 kHz.
 
-The total bandwidth of the downlink sums to 180 kHz.
-That fits neatly into a GSM transmission channel (TCH) which is 200 kHz wide.
-The LTE guard band varies in width, depending on carrier bandwidth.
+NB-IoT devices differ from other mobile devices in that they aren't downlink critical.
+The sensor deployment requires that data from the sensors have to be communicated with other devices or a central service.
+The uplink is therefore the important direction.
+It can be further assumed that the devices initiates the communication, not the base station.
+\hghlght{Investigate how the paging mechanism is implemented in NB-IoT.
+There must be some way for the base station to reach the UE.}
+
+%The LTE guard band varies in width, depending on carrier bandwidth.
 Most deployments of LTE though, use 5 MHz or larger bandwidth, since the smaller bandwidths introduce exceptions from the normal structure.
@@
 That way both functionalities (LTE and NB-IoT) are active at the same time.
+
+\subsection{Channel structure}
+
+\hghlght{It is important to describe the physical channels, since the analysis goes on about identifying when they reach congestion}.
+\point{Needs to be investigated if the higher layer channels introduce any limitation.}
+
+\subsection{Access}
+
+As mentioned, NB-IoT is uplink critical communication meaning that the device initiates the communication and the data exchanged is something the device has been collecting.
+This means that one of the most critical capacity aspects of NB-IoT is how many devices can be accessing the network and how long does it take for the device to ,,get rid of the data''.
+
+The access process is as follows: \hghlght{These steps have to be verified from \cite{3gpp}}
+\begin{enumerate}
+  \item Random access: the device initiates a connection with the base station using the random channel
+  \begin{enumerate}
+    \item Here the possibility of collision with other devices occurs i.e. when two devices try to access at same time they interfere and neither of them gets through
+    \item \hghlght{What is the retry mechanism ?}
+  \end{enumerate}
+  \item The base station assigns resources on the dedicated channels \point{freeing resources from the random access channel ??}
+  \item When the device has gotten its dedicated resources it can begin transmitting
+\end{enumerate}
+
+The capacity limit is reached when the number of devices increases, partly due to the probability of collisions and partly due to the capacity of the dedicated channels.
+\hghlght{The point of the paper is to identify which one is congested before.}
 
@@
 Schlienz (2016) is a really god reference to work the access method from.
-1
 

### 2022-03-20 writing/draft.tex
Progress

@@
 
-According to 3GPP there are three implementations of wide area network (WAN) narrowband internet-of-things (NB-IoT).
-One uses a single GSM transmission channel (TCH) and is integrated into the GSM network.
-The other two are co-existing with LTE where one uses the guard band between LTE channels and the other is embedded into the LTE channel.
+3GPP defines wide area network (WAN) narrowband internet-of-things (NB-IoT) to handle communication with connected devices.
+The physical channel structure is the same as LTE, but scaled down to occupy less bandwidth.
+The downscaling provides flexibility to place the carrier.
+One solution is to place it on a single  GSM transmission channel (TCH), and that way integrate it with the GSM network.
+The other two are co-existing with LTE, where one uses the guard band between LTE channels and the other is embedded into the LTE carrier \cite{schlienz}.
 
-Regardless of where the carrier resides, the physical channel structure is always the same.
-It is based on LTE, but scaled down to occupy less bandwidth which gives an opportunity to place the carrier with greater flexibilty.
-The need for high throughput isn't relevant for NB-IoT, since the devices send limited amount of data.
-The LTE structure in itself incorporates high throughput.
-This can be utilised by NB-IoT, adding further data protection and reconstruction.
-That gives added capability to handle channel path loss of up to 20 dB {[\hghlght{missing reference}}].
+The high throughput characteristics of LTE, and the fact that NB-IoT does not need to send big volumes of data, gives an opportunity to increase the protection of the transmitted data.
+This is done by extra padding, more robust coding scheme and modulation.
+The benefit is increased link budget, allowing up to 20 dB extra attenuation, and decreased probability of wrongly received data, requiring retransmission \cite{wang}.
 
-NB-IoT downlink is defined to use 12 sub-carriers, with each sub-carrier having bandwidth of 15 kHz.
-That adds up to 180 kHz, which matches perfectly the GSM transmission channel (TCH) of 200 kHz.
-The uplink uses either a single sub-carrier, which can be of varying bandwidth: 3,75 kHz or 15 kHz, or it uses 3, 6, or 12 sub-carriers of 15 kHz bandwidth each\cite{3gpp}.
-The last is the same as for the downlink and the total bandwidth becomes 180 kHz.
-
-NB-IoT devices differ from other mobile devices in that they aren't downlink critical.
-The sensor deployment requires that data from the sensors have to be communicated with other devices or a central service.
+NB-IoT devices differ from other mobile devices, in that they aren't downlink critical.
+The sensor deployment requires data to be transmitted from the sensors, towards other devices or a central service.
 The uplink is therefore the important direction.
 It can be further assumed that the devices initiates the communication, not the base station.
-\hghlght{Investigate how the paging mechanism is implemented in NB-IoT.
-There must be some way for the base station to reach the UE.}
+
+An additional design criteria for NB-IoT, is long battery life \cite{wang}.
+Having the extra link budget, it can transmitt with lower power but it is also important to avoid unnecessary retransmission.
+In the case of retransmission, there is waiting - awake - time and extra transmission.
+Both require extra battery usage which has to be avoided.
 
@@
 \subsection{Channel structure}
+
+NB-IoT downlink is defined to use 12 sub-carriers, with each sub-carrier having bandwidth of 15 kHz.
+That adds up to 180 kHz, which fits perfectly into the GSM transmission channel (TCH) of 200 kHz.
+The uplink uses either a single sub-carrier, which can be of varying bandwidth: 3,75 kHz or 15 kHz, or it uses 3, 6, or 12 sub-carriers of 15 kHz bandwidth each\cite{3gpp}.
+The last is the same as for the downlink and the total bandwidth becomes 180 kHz.
 
@@
 \hghlght{The point of the paper is to identify which one is congested before.}
+
+\point{Notes:}\vspace{-4mm}
+\begin{itemize}
+  \item \hghlght{Investigate how the paging mechanism is implemented in NB-IoT. There must be some way for the base station to reach the UE.}
+\end{itemize}
 
@@
 \begin{thebibliography}{4}
+\bibitem{schlienz}
+Schlienz, J., \& Raddino, D. (2016).
+\textit{Narrowband internet of things whitepaper}. White Paper, Rohde\&Schwarz, 1-42.
+
 \bibitem{wang}
@@
 \textit{A primer on 3GPP narrowband Internet of Things}. IEEE communications magazine, 55(3), 117-123.
-
-\bibitem{schlienz}
-Schlienz, J., \& Raddino, D. (2016).
-\textit{Narrowband internet of things whitepaper}. White Paper, Rohde\&Schwarz, 1-42.
 

### 2022-04-26 writing/draft.tex
new approach

@@
 The physical channel structure is the same as LTE, but scaled down to occupy less bandwidth.
-The downscaling provides flexibility to place the carrier.
+The downscaling provides flexibility to place the carrier where there is available space.
 One solution is to place it on a single  GSM transmission channel (TCH), and that way integrate it with the GSM network.
@@
 
-The high throughput characteristics of LTE, and the fact that NB-IoT does not need to send big volumes of data, gives an opportunity to increase the protection of the transmitted data.
+The high throughput characteristics of LTE, and the fact that NB-IoT devices do not need to send big volumes of data, gives an opportunity to increase the protection of the transmitted data.
 This is done by extra padding, more robust coding scheme and modulation.

### 2022-04-26 code/simulation.py

@@ -8,39 +8,30 @@ from numpy import random
 RANDOM_SEED = 42
 RADIO_SIZE = 2
 
+NUM_UE = 1000
+
 TX_BYTES = [0, 100]
 T_INTER = [30, 300]        # Create a car every [min, max] seconds
 SIM_TIME = 1000            # Simulation time in seconds
 
-
-def ue_gen(env, rach):
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+def simEnv(env):
   while True:
-    for it in (random.randint(0,2)):
-      print(env.now, "attempt", it)
-      req = env.process(rach_req(env, rach))
-      yield req
-
-      rach.release(req)
+    for ue in range(NUM_UE):
+      if (random.uniform(1000)<100):
+        print(str(env.now)+":", ue, "transmitted !")
 
     # One step at a time
     yield env.timeout(1)
 
-def rach_req(env, rach):
-  req = rach.request()
 
-  if (req):
-    print (env.now, "avaiable resources")
-  
-    print("Time now:", env.now)
-    yield env.timeout(1)
-  
-  return(req)
-
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
 env = simpy.Environment()
-rach = simpy.Resource(env, 1)
+# rach = simpy.Resource(env, 1)
 
-env.process(ue_gen(env, rach))
-env.run(until=20)
+env.process(simEnv(env))
+env.run(until=200)
 
 """
 def ue(name, env, base_station, radio_res):

### 2022-05-03 code/simulation.py
Working simulation, missing tuning

@@ -2,83 +2,102 @@
 
 import simpy
 import itertools
+import pandas as pd
 from numpy import random
 
-
-RANDOM_SEED = 42
-RADIO_SIZE = 2
-
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+# ENV variables
+# NUM_UE is the number of devices in the simulation environment
 NUM_UE = 1000
 
-TX_BYTES = [0, 100]
-T_INTER = [30, 300]        # Create a car every [min, max] seconds
-SIM_TIME = 1000            # Simulation time in seconds
+# SIM_TIME the number of iteration over the simulation
+# one unit is one millisecond i.e. one TTI - not yet
+SIM_TIME = 10000
+
+# AVG_IDLE is the average time between transmissions
+# It is set to 5 minuttes or 300.000 ms
+AVG_IDLE = 300
+
+# AVG_BYTES is the average number of bytes transmitted per session
+AVG_BYTES = 4e3
+
+# AVG_TX is the average number of bytes transmitted per TTI
+TX_AVG = 1000
+TX_VAR = 20
+
+
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+# MEAS VARIABLES
+progRes = pd.DataFrame(columns=['ts','ue','txBytes','remainBytes','state'])
+
+# UE's generated unconnected with exponentially distributed waiting time
+# the average waiting time is 1 hour i.e. the device sends every hour some data
+# NOTE: one constraint could be minimum 1 hour between transmissions
+ueBase = { i :{'state':'idle', 'nextTx':round(random.exponential(AVG_IDLE)), 'txBytes':0} for i in range(NUM_UE) }
+
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
[…]

### 2022-05-03 writing/draft.tex
Progress

@@
 
-\newcommand{\tilv}[1]{\textbf{\color{ymagreen} [#1]}}
-\newcommand{\note}[1]{\textbf{\color{ymagray}[#1]}}
+\newcommand{\tilv}[1]{\textbf{\color{ymagreen}[#1]}}
+\newcommand{\note}[1]{\textbf{[\color{ymablue}#1}]}
 \newcommand{\highlight}[1]{{\bluen{#1}}}
-\newcommand{\hghlght}[1]{\textbf{\color{ymaorng} #1}}
+\newcommand{\hh}[1]{\textbf{\color{ymaorng} #1}}
 \newcommand{\point}[1]{\textbf{\color{ymared} #1}}
@@
 \clearpage
-\section{NB-IoT}
+\section{NB-IoT}\label{theo}
 
@@
 
-\hghlght{It is important to describe the physical channels, since the analysis goes on about identifying when they reach congestion}.
+\hh{It is important to describe the physical channels, since the analysis goes on about identifying when they reach congestion}.
 \point{Needs to be investigated if the higher layer channels introduce any limitation.}
@@
 
-The access process is as follows: \hghlght{These steps have to be verified from \cite{3gpp}}
+The access process is as follows: \hh{These steps have to be verified from \cite{3gpp}}
 \begin{enumerate}
@@
     \item Here the possibility of collision with other devices occurs i.e. when two devices try to access at same time they interfere and neither of them gets through
-    \item \hghlght{What is the retry mechanism ?}
+    \item \hh{What is the retry mechanism ?}
   \end{enumerate}
@@
 The capacity limit is reached when the number of devices increases, partly due to the probability of collisions and partly due to the capacity of the dedicated channels.
-\hghlght{The point of the paper is to identify which one is congested before.}
+\hh{The point of the paper is to identify which one is congested before.}
 
@@
 \begin{itemize}
-  \item \hghlght{Investigate how the paging mechanism is implemented in NB-IoT. There must be some way for the base station to reach the UE.}
+  \item \hh{Investigate how the paging mechanism is implemented in NB-IoT. There must be some way for the base station to reach the UE.}
 \end{itemize}
@@
 When the user has gotten the random access procedure completed and is scheduled on the transmission channels.
-The next step is the focus of the thesis - how to maximise ??, to be able to service more users and withholding the goal of minimum transmission.w
+The next step is the focus of the thesis - how to maximise ??, to be able to service more users and withholding the goal of minimum transmission.
 
@@
 % -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+% HERMUN
+% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
+\clearpage
+\section{Simulation}
+
+To simulate the environment there are a number of key features that need to be taken into account:
+
+\begin{itemize}
+\item What is the expected number of UE trying to send \note{base this on articles}
+\item How frequent do the UE send like eg. how many times a day \note{base this on articles}
+\item What is the package size that is sent per transmission (full session) \note{base this on articles}
+\item And how long does it take to process the package
+\end{itemize}
+
+The simulation is a Python script relying on the package SimPy.
+This is an event driven simulation package, but currently the events aren't used.
+Instead it is utilised such that one step represents one TTI.
+The script could be extended to have a limited number of RACH channels \note{assuming it is one available per TTI}, thus making use of the built in features as events.
+
+\subsection{Assumptions}
+
+The initial assumption, is to look at an area of one base station (eNodeB).
+According to Ericsson blog \cite{lme_web1} the base station can serve an area up to 40 km and have around 50.000 connections on the cell.
+A more pragmatic solution is to look at a rural area of 20 km radius.
+Such an area can have \hh{xx} homes.
+If a home is expected to have \hh{yy} NB-IoT devices, the scenario gives \hh{xx*yy} connections.
+
+Here the focus is on smart metering devices. A measurement is sent every 5-15 minuttes and the size of the package is \hh{bb}. \note{The package size is random variable with \?\? distribution}
+
+% -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
 % Heimildaskrá
@@
 
+\bibitem{lme_web1}
+Ericsson. (2019).
+\textit{What is NB-IoT? Practical tips to unlock its business potential}.
+Ericsson blog.
+\url{https://www.ericsson.com/en/blog/2019/10/what-is-nb-iot}
+
 \end{thebibliography}
@@
 
-\hghlght{Statement:} For a device with 164 dB coupling loss, a 10-year battery life can be reached if the UE transmits 200-byte data a day on average {[TR 45.820 v13.1.0, “Cellular system support for ultra low complexity and low throughput internet of things,” Nov.
+\hh{Statement:} For a device with 164 dB coupling loss, a 10-year battery life can be reached if the UE transmits 200-byte data a day on average {[TR 45.820 v13.1.0, “Cellular system support for ultra low complexity and low throughput internet of things,” Nov.
 2015]}

### 2022-09-10 writing/draft.tex
Update

@@
 
-NB-IoT downlink is defined to use 12 sub-carriers, with each sub-carrier having bandwidth of 15 kHz.
-That adds up to 180 kHz, which fits perfectly into the GSM transmission channel (TCH) of 200 kHz.
-The uplink uses either a single sub-carrier, which can be of varying bandwidth: 3,75 kHz or 15 kHz, or it uses 3, 6, or 12 sub-carriers of 15 kHz bandwidth each\cite{3gpp}.
-The last is the same as for the downlink and the total bandwidth becomes 180 kHz.
+The most common implementation of NB-IoT in Europe is using Frequency Division Duplexing (FDD) \note{Missing reference}. The following description will therefore only focus on Type 1 frame structure.
+
+One NB-IoT radio downlink frame spans 10 ms i.e.
+the time each device is assigned to receive from the base station (eNB).
+This frame is divided into 10 subframes, where each is two slots.
+One slot consists of 7 Orthogonal Frequency Multiple Access (OFDMA) symbols with varying symbol size.
+
+In the frequency domain the bandwidth is divided into 12 sub-carriers, with each sub-carrier having bandwidth of 15 kHz.
+This gives a total bandwidth of 180 kHz and a grid equivalent to one LTE resource block (RB), as shown in figure \ref{fig:lteRb} \cite{schlienz}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.7\textwidth]{img/The-Resource-Block-in-LTE.png}
+  %\vspace{-0mm}
+  \caption{Resource block structure of LTE}
+  \label{fig:lteRb}
+\end{figure}
+
+The uplink differs from the downlink in that it uses Single Carrier Frequency Division Multiple Access (SC-FDMA).
+It has the same total bandwidth but can be implemented with two types of sub-carrier bandwidth: 15 kHz or 3,75 kHz.
+The bandwidth is decided with configuration in the eNB
+
+In the case of 15 kHz, the grid structure is the same as for the downlink. In the other case
+
+, or it uses 3, 6, or 12 sub-carriers of 15 kHz bandwidth each \cite{3gpp}. The last is the same as for the downlink and the total bandwidth becomes 180 kHz.
+
+\hh{The bandwidth has the advantage of fitting into the GSM transmission channel (TCH) of 200 kHz}.
 
@@
 
-Here the focus is on smart metering devices. A measurement is sent every 5-15 minuttes and the size of the package is \hh{bb}. \note{The package size is random variable with \?\? distribution}
+Here the focus is on smart metering devices. A measurement is sent every 5-15 minuttes and the size of the package is \hh{bb}. \note{The package size is random variable with ?? distribution}
 
